---
title: "Introduction"
author: "Robin Evans"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

## Preliminaries

You'll need a few packages to get the most out of this one.
```
install.packages("rje")
install.packages("BiDAG")
```
The `graph` and `Rgraphviz` packages are hosted on Bioconductor:
```
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install(version = "3.10")
library("graph")
library("Rgraphviz")
```
Then call:
```{r, message=FALSE}
library(DAGtools)
```

# A Basic Demonstration

## Discrete Data

First set up a distribution that comes from the DAG
$X \rightarrow Z \leftarrow Y$.  
```{r get_prob_dist}
p <- c(outer(c(0.4,0.6), c(0.7,0.3)))*c(0.2,0.7,0.4,0.8, 0.8,0.3,0.6,0.2)
dim(p) <- rep(2,3)
```

Now we can generate some data.
```{r get_disc_dat}
set.seed(41)
n <- 1e3  # number of samples
levs <- combinations(rep(2,3))  # from rje package
dat <- levs[sample(1:8, size=n, replace=TRUE, prob = p),]
head(dat)  # first few samples
```

Let's set up the MCMC algorithm.  We can use the wrapper function
`fit_mcmc`.  Specify `scoretype = "bde"` for discrete data, 
"bge" for continuous.
```{r do_mcmc_1, message=FALSE}
out <- fit_mcmc(dat, scoretype = "bde", iterations = 1e5)
out <- summary(out)
plot(out)
```
In the example above, for instance, the edge pointed from $X \rightarrow Z$
in `r round(100*out$adj[1,3])`% of the samples, and $X \leftarrow Z$ in the remainder 
(`r round(100*out$adj[3,1])`%).  In this case it has (erroneously) concluded
that one of the three graphs $X \rightarrow Z \rightarrow Y$ or 
$X \leftarrow Z \rightarrow Y$ or $X \leftarrow Z \leftarrow Y$ generated
the data, and these are indistinguishable and therefore equally likely.

We can increase the sample size, and should obtain the correct graph.
```{r do_mcmc_1a, message=FALSE, cache=TRUE}
n <- 1e4  # number of samples
dat <- levs[sample(1:8, size=n, replace=TRUE, prob = p),]
out <- fit_mcmc(dat, scoretype = "bde", iterations = 1e5)
out2 <- summary(out)
out2
```
This time we correctly obtain $X \rightarrow Z$
in `r round(100*out2$adj[1,3],1)`% of the samples (and $Y \rightarrow Z$ in 
`r round(100*out2$adj[2,3],1)`%).

### Sampled Graphs

The algorithm provides a sequence of separate sampled DAGs, which we 
can individually inspect.  
```{r incidence}
out2$incidence[[1]]
```
The `plot` function can be used to give a summary of the results (by
default, it shows edges present in 20% of graphs); a double-headed 
arrow means both directions were present at least this often.  
```{r plot_graph}
plot(out2)
```

## Gaussian Data

Suppose we have continuous data instead, and wish to use 
a Gaussian likelihood.
```{r covariance}
set.seed(42)
n <- 1e4  # number of samples
B <- diag(4)
B[1,3] = 0.3; B[1,4] = 0.2   # regression coefficients
B[2,3] = 0.25; B[3,4] = 0.5 
dat <- matrix(rnorm(4*n), ncol=4) %*% solve(B)

cov(dat)
```
You can use the function `fit_mcmc` to run things behind 
the scenes:
```{r do_mcmc2}
out <- fit_mcmc(dat, scoretype = "bge", iterations = 1e5)
out2 <- summary(out)
out2
```
For Gaussian data, we can augment the sampled graphs 
with parameter samples.  
```{r augment}
out2 <- augment_mcmc(out2, dat, verbose = FALSE)
names(out2)
```
Once augmented, we can inspect the estimated 
causal effects:
```{r hist}
hist(out2$CauEff[2,3,], breaks=50)
```
The plot method for \code{MCMC_summary} variables 

```{r plot_graph2}
plot(out2)
```

## Mediation

Using Gaussian data, we can compute mediators:
```{r}
cau_path_34 <- sum_causal_paths(out2, 3, 4)
```
We can also plot causal effects with an augmented chain:
```{r}
plot_causal_effects(out2)
```

## Multiple Data Sets

```{r fit_multiple}
set.seed(47)
n <- 1e4
dat2 <- matrix(rnorm(4*n), ncol=4) %*% solve(B)
out <- fit_multiple(list(dat, dat2), iterations=1e3)
```
This returns an augmented chain, so we can (for example)
try a plot of the causal effects:
```{r plot_graph3}
plot_causal_effects(out)
```
